Concurrent Programming with CXL

### 2020-05-21 book/paper.tex
Stuff about invariant checking

To see that invariant $I_1$ is maintained, note that
\texttt{\^{}p == True} upon entry of \texttt{tas}.
So there are two cases:
\begin{enumerate}
\item \texttt{\^{}s} is \texttt{False} upon entry to \texttt{tas}.
Then upon exit \texttt{\^{}p == False} and \texttt{\^{}s == True}, maintaining
the invariant.
\item \texttt{\^{}s} is \texttt{True} upon entry to \texttt{tas}.
Then upon exit nothing has changed, maintaining the invariant.
\end{enumerate}
Invariant $I_1$ is also easy to verify for exiting the critical section.

@@

Invariant $I_2$ is obvious as process $i$ only proceeds to the critical
section if \texttt{private[$i$] == False} and no other process modifies
\texttt{private[$i$]}.
Indeed, there is no need to keep \texttt{private} as a shared
variable---we only did so to be able to show the invariants.
Figure~\ref{fig:tas2} shows a more straightforward implementation.
Compared to Peterson's Algorithm, this approach is general for
any number of processes.

>>>>

To see that invariant $I_1$ is maintained, note that
\texttt{\^{}p == True} upon entry of \texttt{tas}.
So there are two cases:
\begin{enumerate}
\item \texttt{\^{}s} is \texttt{False} upon entry to \texttt{tas}.
Then upon exit \texttt{\^{}p == False} and \texttt{\^{}s == True}, maintaining
the invariant.
\item \texttt{\^{}s} is \texttt{True} upon entry to \texttt{tas}.
Then upon exit nothing has changed, maintaining the invariant.
\end{enumerate}
Invariant $I_1$ is also easy to verify for exiting the critical section.
Invariant $I_2$ is obvious as process $i$ only proceeds to the critical
section if \texttt{private[$i$] == False} and no other process modifies
\texttt{private[$i$]}.

\begin{figure}
\begin{code}
\verbatiminput{UpTASinv.cxl}
\end{code}
\caption{Checking invariants.}
\label{fig:tasinv}
\end{figure}

CXL can check these invariants as well.  Figure~\ref{fig:tas} already
has the code to check $I_2$.  But how would one go about checking an
invariant like $I_1$.  Invariants must hold for every state.
For $I_2$ we only need an assertion at label \texttt{@cs} because the
premisse is that there is a process at that label.  However, we would
like to check $I_1$ in \emph{every state} (after the variables have
been initialized).

We can do this by adding another process that, in a loop,
checks the invariant.  Figure~\ref{fig:tasinv} shows the code.
Method \texttt{checkInvariant()} checks to see if the invariant holds
in a state.  It introduces a new feature of CXL: the ability to have
variables local to a method.  In this case, the variable \texttt{sum}
is used to compute the number of variables that have value
\texttt{False}.
The function is invoked by in a loop by a process that runs alongside
the other processes.
Note that this process must be started after the variables are
initialized---before that the invariant does not hold yet.
Because CXL tries every possible execution, the process is guaranteed
to find violations of the invariant if it does not hold.

@@

Because each \texttt{private[i]} is only access by process~$i$,
there is no need to keep \texttt{private} as a shared
variable---we only did so to be able to show and check the invariants.
Figure~\ref{fig:tas2} shows a more straightforward implementation.
Compared to Peterson's Algorithm, this approach is general for
any number of processes.
